Testing production scenarios for (anti-)(hyper-)nuclei with multiplicity-dependent measurements at the LHC

### 2019-03-31 EPIPHANY2019proceedings/Bellini_Epiphany2019_proc.tex
Continuing with models

For our study, we consider nuclei and hyper-nuclei with mass number $A = 2, 3$ and 4, whose properties are summarised in Tab. \ref{tab:nucleusradii}. Those properties are the same as for their anti-matter counterparts and we assume that the same formation mechanisms are at play for matter and anti-matter\footnote{For brevity, in the following we refer to ``nuclei'' and ``hyper-nuclei'' but we imply both matter and anti-matter.}.
Nuclei and hyper-nuclei are special objects with respect to non-composite hadrons (pions, protons, etc.), because their size is comparable to a fraction or the whole system created in pp, \pPb~and \PbPb~collisions.
The size is typically defined in two ways: a) as the rms of the (charge) distribution ($\lambda_A$), typically measured in electron scattering experiments, or b) as the size parameter of the object wave-function ($r_A$), typically taken as the gaussian solution of an isotropic harmonic oscillator potential in coalescence models. 
For light nuclei, $\lambda_A\approx$~2~fm.  For the hyper-triton, theoretical calculations indicate a charge rms radius $\lambda_A\approx$~5~fm \cite{Nemura:1999qp}, driven by the average separation of the $\Lambda$ relative to the two other nucleons. 
Assuming a similar structure (e.g. a s-wave interaction for a bound state of a n or a $\Lambda$ with a
deuteron), the hypertriton results in a much larger object than the other non-strange nuclei with $A=3$.
A simple relation holds between $\lambda_A$ and $r_A$, see \cite{Bellini:2018epz}. 
In Tab.~\ref{tab:nucleusradii} the binding energy ($B_E$) is also reported. The most tightly bound nucleus is \hefour, whereas the most loosely bound object is \hthreelambda, that is also the largest one. For the latter, we report the separation energy of the $\Lambda$ baryon from the deuteron ($B_\Lambda=0.13$~MeV). 
The large size and the low binding energy of the \hthreelambda~with respect to the other (hyper-)nuclei has important consequences on its production, as discussed in what follows.

\begin{table}[htb]
\resizebox{\textwidth}{!}{%
\centering
\begin{tabularx}{1.25\textwidth}{cccccccc}
\hline \hline \\[-2ex]
\makecell{Mass \\number } & Nucleus           &  \makecell{Compo-\\sition}               & $B_{E}$ (MeV)   & \makecell{Spin \\$J_{A}$} & \makecell{(Charge) \\rms radius \\ \rmsradius$^{meas}$~(fm)} &  \makecell{Harmonic \\ oscillator \\ size \\ parameter \\$r_{A}$ (fm) } & Refs. \\[1ex]  \hline \\[-2ex] 
      A = 2                     & d                                    & pn                                  &   2.224575 (9)     &     1   & 2.1413 $\pm$ 0.0025      &  3.2    &   \cite{VanDerLeun:1982bhg,Mohr:2015ccw}     \\[0.5ex]  \hline \\[-2ex]
\multirow{3}{*}{A = 3}  & \tritium 	                  & pnn                               &    8.4817986 (20) & 1/2   &  1.755  $\pm$ 0.086        &  2.15   &   \cite{Purcell:2015gtm}           \\
                                   & \hethree                         & ppn                                &   7.7180428  (23) & 1/2  & 1.959 $\pm$  0.030         &   2.48  &   \cite{Purcell:2015gtm} \\
                                   & \hthreelambda               & p$\Lambda$n                &    0.13 $\pm$ 0.05 & 1/2  &  4.9 --  10.0                    &  6.8 -- 14.1 & \cite{Davis:2005mb,Nemura:1999qp} \\[0.5ex]  \hline \\[-2ex]
\multirow{4}{*}{A = 4}  & \hefour                          & ppnn                              &    28.29566   (20)  &      0  &  1.6755 $\pm$ 0.0028  &  1.9  & \cite{1674-1137-41-3-030003,Angeli:2013epw} \\
                                   & \hfourlambda                & p$\Lambda$nn              &  2.04 $\pm$0.04   &   0   &    2.0 -- 3.8             & 2.4 -- 4.9  & \cite{Davis:2005mb,Nemura:1999qp} \\
                                   & \hfourtwolambda          &  p$\Lambda\Lambda$n &   0.39 -- 0.51         &    1     &    4.2 -- 7.1          & 5.5 -- 9.4  &   \cite{Nemura:1999qp} \\
                                   & \hefourlambda              & pp$\Lambda$n              &  2.39 $\pm$ 0.03  &    0   &    2.0 -- 3.8            & 2.4 -- 4.9  & \cite{Davis:2005mb,Nemura:1999qp}\\[0.5ex]  \hline \hline
\end{tabularx}
}
\caption{Properties of nuclei and hyper-nuclei with mass number $A \leq 4$. $B_{E}$ is the binding energy in MeV.   The size parameter $r_{A}$, is chosen to approximately reproduce the measured/expected rms,  \rmsradius$^{meas}$~(fm). The proton rms charge radius $\lambda_{p} = 0.879(8)$ fm \cite{bernauer10} is subtracted quadratically from the measured rms charge radius $\rmsradius^{meas}$ of the nucleus $\rmsradius = \sqrt{(\rmsradius^{meas})^2 - {\lambda_{p}^2}}$ to account for the finite extension of the constituents. Implicitly we assume here that $\lambda_{\Lambda}\approx \lambda_{n}\approx \lambda_{p}$. The spin of \hfourtwolambda\ is discussed in the text of~\cite{Nemura:1999qp}.}
\label{tab:nucleusradii}
\end{table}

\subsection{The coalescence approach}

Starting from the model described in~\cite{Scheibl:1998tk, Blum:2017qnn}, we have obtained in \cite{Bellini:2018epz} a generalised expression for the coalescence parameter $B_A$

\begin{equation}\label{eq:master}
	\boxed {  B_{A} = {2J_{A} + 1 \over 2^{A}} {1 \over \sqrt{A}} {1 \over m_{T}^{A-1}} \Bigl({2\pi \over R^2 + ({r_A \over 2})^2 }\Bigr)^{\frac{3}{2}(A-1)} } \;\;,
\end{equation}

\noindent which is a function of the spin of the particle $J_A$, its transverse mass $m_{\rm T}$, its size parameter $r_A$ and the source radius $R$.
Figure \ref{Fig:BA} shows the source radius dependence of $B_A$ for different composite objects, including the nuclei and hyper-nuclei with $A = 2, 3$ and 4 whose properties are reported in Tab.~\ref{tab:nucleusradii}.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.8\textwidth]{coalescenceBA075.eps}
\caption{(Color online) Coalescence parameter $B_A$ as a function of the source radius $R$ as predicted from the coalescence model (Eq.~\ref{eq:master}) for various composite objects with \pt$/A$~=~0.75~\GeVc. For each (hyper-)nucleus, the radius $r$ used for the calculation is reported in the legend.}
\label{Fig:BA}
\end{center}
\end{figure}

\subsection{Thermal-statistical models} 
Thermal-statistical models ~\cite{Andronic:2010qu, Andronic:2017} have been successful in describing the production of light \mbox{(anti-)}(hyper-)nuclei across a wide range of energies in AA collisions, including production at the LHC.
In this approach, particles are produced from a fireball in thermal equilibrium with temperatures of $T_{chem} \approx$ 156 MeV.

>>>>

For our study, we consider nuclei and hyper-nuclei with mass number $A = 2, 3$ and 4, whose properties are summarised in Tab. \ref{tab:nucleusradii}. Those properties are the same as for their anti-matter counterparts and we assume that the same formation mechanisms are at play for matter and anti-matter\footnote{For brevity, in the following we refer to ``nuclei'' and ``hyper-nuclei'' but we imply both matter and anti-matter.}.
Nuclei and hyper-nuclei are special objects with respect to non-composite hadrons (pions, protons, etc.), because their size is comparable to a fraction or the whole system created in pp, \pPb~and \PbPb~collisions.
The size is typically defined in two ways: a) as the rms of the (charge) distribution ($\lambda_A$), typically measured in electron scattering experiments, or b) as the size parameter of the object wave-function ($r_A$), typically taken as the gaussian solution of an isotropic harmonic oscillator potential in coalescence calculations. 
For light nuclei, $\lambda_A\approx$~2~fm.  For the hyper-triton, theoretical calculations indicate a charge rms radius $\lambda_A\approx$~5~fm \cite{Nemura:1999qp}, driven by the average separation of the $\Lambda$ relative to the two other nucleons. 
Assuming a similar structure (e.g. a s-wave interaction for a bound state of a n or a $\Lambda$ with a
deuteron), the hypertriton results in a much larger object than the other non-strange nuclei with $A=3$.
A simple relation holds between $\lambda_A$ and $r_A$, see \cite{Bellini:2018epz}. 
In Tab.~\ref{tab:nucleusradii} the binding energy ($B_E$) is also reported. The most tightly bound nucleus is \hefour, whereas the most loosely bound object is \hthreelambda, that is also the largest one. For the latter, we report the separation energy of the $\Lambda$ baryon from the deuteron ($B_\Lambda=0.13$~MeV). 
The large size and the low binding energy of the \hthreelambda~with respect to the other (hyper-)nuclei has important consequences on its production, as discussed in what follows.

\begin{table}[htb]
\resizebox{\textwidth}{!}{%
\centering
\begin{tabularx}{1.25\textwidth}{cccccccc}
\hline \hline \\[-2ex]
\makecell{Mass \\number } & Nucleus           &  \makecell{Compo-\\sition}               & $B_{E}$ (MeV)   & \makecell{Spin \\$J_{A}$} & \makecell{(Charge) \\rms radius \\ \rmsradius$^{meas}$~(fm)} &  \makecell{Harmonic \\ oscillator \\ size \\ parameter \\$r_{A}$ (fm) } & Refs. \\[1ex]  \hline \\[-2ex] 
      A = 2                     & d                                    & pn                                  &   2.224575 (9)     &     1   & 2.1413 $\pm$ 0.0025      &  3.2    &   \cite{VanDerLeun:1982bhg,Mohr:2015ccw}     \\[0.5ex]  \hline \\[-2ex]
\multirow{3}{*}{A = 3}  & \tritium 	                  & pnn                               &    8.4817986 (20) & 1/2   &  1.755  $\pm$ 0.086        &  2.15   &   \cite{Purcell:2015gtm}           \\
                                   & \hethree                         & ppn                                &   7.7180428  (23) & 1/2  & 1.959 $\pm$  0.030         &   2.48  &   \cite{Purcell:2015gtm} \\
                                   & \hthreelambda               & p$\Lambda$n                &    0.13 $\pm$ 0.05 & 1/2  &  4.9 --  10.0                    &  6.8 -- 14.1 & \cite{Davis:2005mb,Nemura:1999qp} \\[0.5ex]  \hline \\[-2ex]
\multirow{4}{*}{A = 4}  & \hefour                          & ppnn                              &    28.29566   (20)  &      0  &  1.6755 $\pm$ 0.0028  &  1.9  & \cite{1674-1137-41-3-030003,Angeli:2013epw} \\
                                   & \hfourlambda                & p$\Lambda$nn              &  2.04 $\pm$0.04   &   0   &    2.0 -- 3.8             & 2.4 -- 4.9  & \cite{Davis:2005mb,Nemura:1999qp} \\
                                   & \hfourtwolambda          &  p$\Lambda\Lambda$n &   0.39 -- 0.51         &    1     &    4.2 -- 7.1          & 5.5 -- 9.4  &   \cite{Nemura:1999qp} \\
                                   & \hefourlambda              & pp$\Lambda$n              &  2.39 $\pm$ 0.03  &    0   &    2.0 -- 3.8            & 2.4 -- 4.9  & \cite{Davis:2005mb,Nemura:1999qp}\\[0.5ex]  \hline \hline
\end{tabularx}
}
\caption{Properties of nuclei and hyper-nuclei with mass number $A \leq 4$. $B_{E}$ is the binding energy in MeV.   The size parameter $r_{A}$, is chosen to approximately reproduce the measured/expected rms,  \rmsradius$^{meas}$~(fm). The proton rms charge radius $\lambda_{p} = 0.879(8)$ fm is subtracted quadratically from the measured rms charge radius $\rmsradius^{meas}$ of the nucleus $\rmsradius = \sqrt{(\rmsradius^{meas})^2 - {\lambda_{p}^2}}$ to account for the finite extension of the constituents. Implicitly we assume here that $\lambda_{\Lambda}\approx \lambda_{n}\approx \lambda_{p}$.}
\label{tab:nucleusradii}
\end{table}

\subsection{The coalescence approach}
In the coalescence picture, nucleons produced in the collision coalesce into nuclei if they are close in space and have similar velocities \cite{Butler:1963, Kapusta:1980, Sato:1981ez, Nagle:1996vp, Scheibl:1998tk}.
The coalescence probability is encoded in the coalescence parameter, $B_{A}$.
Considering that at LHC energies the number of produced protons and neutrons at midrapidity as well as their momentum distributions are expected to be equal, $B_{A}$ is defined as
%
%
\begin{equation}
E_{A}\frac{\mathrm{d}^{3}N_{A}}{\mathrm{d}p_{A}^{3}}=B_{A}{\left(E_{\mathrm{p,n}}\frac{\mathrm{d}^{3}N_{\mathrm{p,n}}}{\mathrm{d}p_{\mathrm{p,n}}^{3}}\right)^{A}}\left\vert_{\vec{p}_{\mathrm{p}}=\vec{p}_{\mathrm{n}}=\frac{\vec{p}_{A}}{A}} \right.,
\label{eq:BA}
\end{equation}
%
%
where $p_{\mathrm{p,n}}$ are the proton and neutron momenta and $E_{p,n}$ their energies. 
Equation~\ref{eq:BA} represents also the operative definition of $B_A$ that is used by experiments like ALICE to extract the coalescence probability starting from the measured nucleus and nucleon (proton) distributions.
Starting from the model described in~\cite{Scheibl:1998tk, Blum:2017qnn}, we have obtained in \cite{Bellini:2018epz} a generalised expression for $B_A$
%
%
\begin{equation}\label{eq:master}
B_{A} = {2J_{A} + 1 \over 2^{A}} {1 \over \sqrt{A}} {1 \over m_{T}^{A-1}} \Bigl({2\pi \over R^2 + ({r_A \over 2})^2 }\Bigr)^{\frac{3}{2}(A-1)} \;\;,
\end{equation}
%
%
\noindent which is a function of the spin of the particle $J_A$, its transverse mass $m_{\rm T}$, its size parameter $r_A$ and the source radius $R$. Very importantly, Eq.~\ref{eq:master} takes explicitly into account the source size ($R$), as the coalescence probability naturally decreases for nucleons with similar momenta that are produced far apart in configuration space. 
Moreover, the source size is identified with the  effective sub-volume of the whole system that is governed by the (momentum-dependent) homogeneity length of the interacting nucleons and experimentally accessible with Hanbury-Brown-Twiss (HBT) interferometry~\cite{Scheibl:1998tk, Blum:2017qnn}. 
Figure~\ref{Fig:BA} shows the source radius dependence of $B_A$ for nuclei and hyper-nuclei with $A = 2, 3$ and 4 whose properties are reported in Tab.~\ref{tab:nucleusradii}. 
For the cases in which more than one estimate for $r_A$ is available, we have adopted the lowest value for the calculations in Fig.~\ref{Fig:BA}.
%
%
\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.8\textwidth]{coalescenceBA075.eps}
\caption{(Color online) Coalescence parameter $B_A$ as a function of the source radius $R$ as predicted from the coalescence model (Eq.~\ref{eq:master}) for various composite objects with \pt$/A$~=~0.75~\GeVc. For each (hyper-)nucleus, the radius $r$ used for the calculation is reported in the legend.}
\label{Fig:BA}
\end{center}
\end{figure} 
%
%
We observe that the coalescence probability decreases with increasing mass number and $B_A$ decreases with increasing volume. 
For a given $A$, the larger the object radius, the lower is $B_A$, as clearly visible by comparing \hethree~and \hthreelambda. 
For objects with same $A$, mass and spin (e.g. the isobars \tritium~and \hethree), $B_A$ differs only due to the different radius $r_A$. 
This difference is more relevant in small systems, because in large systems the difference between nucleus radii is much smaller than the size of the source.

The production of objects with radius larger than the source is strongly suppressed, indicating that the process is driven by the lenght scale defined by the object radius relative to the source radius.

\subsection{Thermal + blast-wave model} 
Thermal-statistical models ~\cite{Andronic:2010qu, Andronic:2017} have been successful in describing the production of light \mbox{(anti-)}(hyper-)nuclei across a wide range of energies in AA collisions, including production at the LHC.
In this approach, particles are produced from a fireball in thermal equilibrium with temperatures of $T_{chem} \approx$ 156 MeV.
The thermal model cannot -- alone -- be compared to the \pt-dependent coalescence description because it provides only predictions for \pt-integrated yields. 
A thermal particle production implies a Boltzmann distribution of the momenta only for a static source, which is not the case of the rapidly expanding system produced in heavy-ion collisions. 
The thermal model (i.e. the GSI-Heidelberg implementation we have considered) needs to be complemented by a
hydrodynamic description of the rapidly expanding source. 
To that end, we chose to use the Blast-Wave model, which has been proven to describe reasonably well the measured momentum distributions of protons. 
Our ``thermal+blast-wave'' approach results from the combination of the two models: the \pt-dependence is modelled by the blast-wave whereas the normalisation (i.e. the \pt-integrated yield) is taken from the thermal model predictions.
